\class : Problem Set \psetnum

### 2022-05-16 531_stars/pset4/solutions.tex
fix some typos

\question{Part b - Terminal wind velocity latitude trends}
\answer{
    In rapidly rotating stars, the effective surface gravity at a given radius becomes a function of latitude, with stronger gravity at the poles than the equation. The escape velocity is a direct function of this surface gravity. We have previously shown that the terminal wind velocity is approximately equal to the escape velocity (see Section 15.2.2 of the textbook). For these reasons the terminal wind velocity is higher at the poles than at the equator.
}

\question{Part c - Density of bi-stability disk}
\answer{
    The density of the gas scales as
    \begin{equation}
        \rho \sim \frac{\dot{M}}{v_{\infty}}
    \end{equation}
    For a rotation-induced bi-stability disk, the mass loss is higher than at the poles at its slightly lower $T_{\rm eff}$ places it just below a bi-stability jump (see Section 15.3). Moreover, the terminal wind velocity ($v_\infty$) is higher on the lower side of a bi-stability jump (see Section 15.2.2). The combination of these effects leads to a significant increase in density in the disk relative to the poles.
}

\question{\textbf{27.1 - Limit for NS vs.\ BH for rotating stars}}
\answer{
    The limit of stars with initial masses resulting in NSs or BHs could change due to rotation in the following ways.

    \begin{itemize}
        \item \textit{Case for higher mass limit:} The mass needed to create a BH would need to increase to compensate for the increased mass loss due to rotation. Otherwise, less mass is available to go into the creation of a remnant and so stars that would have resulted in BHs instead result in NSs.
        \item \textit{Case for lower mass limit:} Though the overall mass is decreased for a rapidly-rotating star, the mass of the helium core \textit{increases} due to the induced mixing. A larger core mass leads to a larger remnant mass and thus more easily create a BH with a given initial mass.
    \end{itemize}
    Overall, the winner here is the latter reasoning. Core mass is the most important parameter for determining the destiny of any massive star and so if all stars were rapidly-rotating the mass limit between NSs and BHs would decrease.
}

>>>>

\question{Part b - Terminal wind velocity latitude trends}
\answer{
    In rapidly rotating stars, the effective surface gravity at a given radius becomes a function of latitude, with stronger gravity at the poles than the equator. The escape velocity is a direct function of this surface gravity. We have previously shown that the terminal wind velocity is approximately equal to the escape velocity (see Section 15.2.2 of the textbook). For these reasons the terminal wind velocity is higher at the poles than at the equator.
}

\question{Part c - Density of bi-stability disk}
\answer{
    The density of the gas scales as
    \begin{equation}
        \rho \sim \frac{\dot{M}}{v_{\infty}}
    \end{equation}
    For a rotation-induced bi-stability disk, the mass loss is higher than at the poles at its slightly lower $T_{\rm eff}$ places it just below a bi-stability jump (see Section 15.3). Moreover, the terminal wind velocity ($v_\infty$) is lower on the lower side of a bi-stability jump (see Section 15.2.2). The combination of these effects leads to a significant increase in density in the disk relative to the poles.
}

\question{\textbf{27.1 - Limit for NS vs.\ BH for rotating stars}}
\answer{
    The limit of stars with initial masses resulting in NSs or BHs could change due to rotation in the following ways:
    \begin{itemize}
        \item \textit{Case for higher mass limit:} The mass needed to create a BH would need to increase to compensate for the increased mass loss due to rotation. Otherwise, less mass is available to go into the creation of a remnant and so stars that would have resulted in BHs instead result in NSs.
        \item \textit{Case for lower mass limit:} Though the overall mass is decreased for a rapidly-rotating star, the mass of the helium core \textit{increases} due to the induced mixing. A larger core mass leads to a larger remnant mass and thus more easily create a BH with a given initial mass.
    \end{itemize}
    Overall, the winner here is the latter reasoning. Core mass is the most important parameter for determining the destiny of any massive star and so if all stars were rapidly-rotating the mass limit between NSs and BHs would decrease.
}
